---
title: "HS614 - Exploration and Modelling of MIMIC dataset (Acute MI)"
author: "Prabha Nayak"
date: "5/16/2019"
output: html_document
---

Exploratory Data Analysis, Data Cleaning and visualization:


```{r}
#install.packages('tidyverse')
#install.packages("Amelia")
#install.packages("caret")
#install.packages("rpart.plot")
#install.packages("corrplot")
#install.packages("pscl")
#install.packages("ROCR")
#install.packages("ROSE")
# install.packages("e1071")
# install.packages("randomForrest")


library(dplyr)
library(tidyverse)
library(Amelia)
library(ggplot2)
library(caret)
library(rpart.plot)
library(corrplot)
library(pscl)
library(ROCR)
library(ROSE)
library(e1071)
library(randomForest)

setwd("C:/Users/PG/Desktop/USF/Spring2019/HS614")
heart_df <- read.csv("mimic_final.csv", header = TRUE, sep = ',')

dim(heart_df)

summary(heart_df)
```

Finding the total number of NAs in each column
```{r}
sum(is.na(heart_df$meanbp))
sum(is.na(heart_df$resprate))
sum(is.na(heart_df$heartrate_max))
sum(is.na(heart_df$chloride))
sum(is.na(heart_df$CK))
sum(is.na(heart_df$creatinine))
sum(is.na(heart_df$glucose))
sum(is.na(heart_df$potassium))
sum(is.na(heart_df$sodium))
```
Dropping columns X and religion as it is not required for the study
```{r}
heart_df <- heart_df[, -c(1,4)]
```

Checkingfor number of levels in the ethnicity column
number of levels = 22
This is too many levels for the study and will be collapsed to 5 levels
```{r}
unique(heart_df$ethnicity)

levels(heart_df$ethnicity) <- list(White= c("WHITE", "WHITE - EASTERN EUROPEAN", "WHITE - RUSSIAN", "WHITE - OTHER EUROPEAN"), Other= c("UNABLE TO OBTAIN", "OTHER", "UNKNOWN/NOT SPECIFIED","AMERICAN INDIAN/ALASKA NATIVE", "MULTI RACE ETHNICITY", "PATIENT DECLINED TO ANSWER"), Asian= c("ASIAN - ASIAN INDIAN", "ASIAN", "ASIAN - CHINESE", "ASIAN - OTHER", "ASIAN - FILIPINO", "ASIAN - CAMBODIAN"), Hispanic= c("HISPANIC OR LATINO", "HISPANIC/LATINO - PUERTO RICAN", "HISPANIC/LATINO - DOMINICAN", "PORTUGUESE"), Black= c("BLACK/AFRICAN AMERICAN", "BLACK/CAPE VERDEAN"))

unique(heart_df$ethnicity)
```

Converting age above 300 to 89
```{r}
heart_df$age <- ifelse(heart_df$age >= 300, 89, heart_df$age)
summary(heart_df$age)
```

Limiting CK value to 20000
```{r}
heart_df$CK <- ifelse(heart_df$CK > 20000, NA, heart_df$CK)
summary(heart_df$CK)
```

limiting systolic bp to 350
```{r}
heart_df$sysbp <- ifelse(heart_df$sysbp >350, NA, heart_df$sysbp)
summary(heart_df$sysbp)
```

limiting diastolic BP to 350
```{r}
heart_df$diasbp <- ifelse(heart_df$diasbp >350, NA, heart_df$diasbp)
summary(heart_df$diasbp)
```

limiting Blood oxygen saturation to 100
```{r}
heart_df$spo2 <- ifelse(heart_df$spo2 >100, NA, heart_df$spo2)
summary(heart_df$spo2)
```

limiting max heart rate to 459
```{r}
heart_df$heartrate_max <- ifelse(heart_df$heartrate_max >450, NA, heart_df$heartrate_max)
summary(heart_df$heartrate_max)
```

limiting respiratory rate to 200
```{r}
heart_df$resprate <- ifelse(heart_df$resprate > 200, NA, heart_df$resprate)
summary(heart_df$resprate)
```

##Number of NA's in the CK column = 340
imputing missing values in column CK to remove all NA's 
```{r}
heart_df$CK[is.na(heart_df$CK)] <- mean(heart_df$CK,na.rm=T)
summary(heart_df$CK)
```


Finding correlation between variables
```{r}
correlations <- cor(heart_df[-(1:2)], use= "pairwise.complete.obs", method= "spearman")
corrplot(correlations, method="color")
```

Using na.omit to remove all rows with NA's in the dataframe
```{r}
heart_df2 <- na.omit(heart_df)
```

```{r}
plot(heart_df2$hos_expire_flag, heart_df2$heartrate_max, main="Scatterplot", 
     xlab="expired", ylab="heart rate ", pch=19)
```

creating plots that will show relations between variables
```{r}
a <- ggplot(data = heart_df2, aes( y=sodium,x=chloride))
a + geom_boxplot(col="blue") +  
              ggtitle("Relation of CK and respiratory rate") +
              ylab("sodium")+
              xlab("chloride")
```

Creating the test train split using the createDataPartition function from the caret package. 
A 75:25 split is created
The 
```{r}
set.seed(1305)
heart_df3_train <- createDataPartition(y = heart_df2$hos_expire_flag, p=0.75, list = FALSE)
train <- heart_df2[heart_df3_train, -c(2:3)]
test <- heart_df2[-heart_df3_train, -c(2:3)]
test$hos_expire_flag= factor(test$hos_expire_flag)
dim(train); dim(test)
```

fitting a model with hos_expiry_flag as a function of all other variable to check
```{r}
logfit1 <- glm(hos_expire_flag~., data = train, family = "binomial")
summary(logfit1)

logfit2 <- glm(hos_expire_flag ~ heartrate_max + resprate + spo2 + chloride + CK + creatinine + glucose + potassium + sysbp, data = train, family = "binomial")
summary(logfit2)

logfit3 <- glm(hos_expire_flag ~ heartrate_max + resprate + spo2 + chloride + CK + creatinine + glucose + potassium, data = train, family = "binomial")
summary(logfit3)

plot(logfit3, 1)
```
Interpretation of fitted coefficients
For every mg/dL increase in creatinine level the odds of dying are on average multiplied by 0.216, when all other variables in the model are held constant

For every unit per liter increase in creatinine kinase level the odds of dying are on average multiplied by 0.000293, when all other variables in the model are held constant

For every bpm increase in maximum heartrate the odds of dying are on average multiplied by 0.026, when all other variables in the model are held constant

For every percentge increase in blood oxygen saturation the odds of dying are on average multiplied by -0.14, when all other variables in the model are held constant

Plotting a scatter plot 
```{r}
plot(train$hos_expire_flag, train$heartrate_max, main="Scatterplot", 
     xlab="expired", ylab="heart rate", pch=19)
```

```{r}
pR2(logfit3)

anova(logfit3, test = "Chisq")
```

```{r}
pred.train <- predict(logfit3, newdata=train)

roc.curve(train$hos_expire_flag, pred.train, 
          main="ROC curve")

pred.test <- predict(logfit3, newdata=test)
pred.test

roc.curve(test$hos_expire_flag, pred.test, add=TRUE, col=2, 
          lwd=2, lty=2)

```

Support vector machine

trainControl() before train() to control the comutational naunces of the train() method

```{r}
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

svm_Linear <- train(hos_expire_flag ~., data = train, method = "svmLinear",
                    trControl=trctrl,
                    preProcess = c("center", "scale"),
                    tuneLength = 10)

svm_Linear
```

predicting the test set and building the ROC curve for the predicted variable. An AUC of 84.5% is obtained. 
```{r}

pred_test <- predict(svm_Linear, newdata = test)
pred_test

roc.curve(test$hos_expire_flag, pred_test, 
          main="ROC curve")

#confusionMatrix(pred_test, factor(test$hos_expire_flag))
```

c=1, gamma=1, kernel=linear
```{r}
svmfit_radial = svm(hos_expire_flag~., data = train, kernel = "linear", gamma = 1, cost = 1)
summary(svmfit_radial)
```

c=1, gamma=0.5, kernel=radial
```{r}
svm_radial = svm(hos_expire_flag~. , data = train, kernel = "radial", gamma = 0.5, cost = 1)
summary(svm_radial)
```

This looks like a better fit which is not or underfitting the train data
```{r}
svmfit_poly = svm(hos_expire_flag~., data = train, kernel = "polynomial", degree= 5, cost = 1)
summary(svmfit_poly)
```



Decision tree
```{r}
set.seed(435)
dtree_fit <- train(hos_expire_flag ~. , data = train, method = "rpart", parms = list(split = "information"), trControl=trctrl, tuneLength = 10)

dtree_fit
```


```{r}
pred_test1 <- predict(dtree_fit, newdata = test)
pred_test1

roc.curve(test$hos_expire_flag, pred_test1, 
          main="ROC curve")

#confusionMatrix(pred_test, factor(test$hos_expire_flag))
```

Random Forrest
```{r}

mi.rf <- randomForest(hos_expire_flag ~ ., data=train, importance=TRUE,
                        proximity=TRUE)
print(mi.rf)

round(importance(mi.rf), 2)

varImpPlot (mi.rf)
```









